---
title: "PA1_template.Rmd"
author: "Will S."
date: "3/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This markdown file would explain the code and reasoning of the code.
###Question 1. Loading and preprocessing the data
Show any code that is needed to:
*Load the data (i.e. read.csv())
*Process/transform the data (if necessary) into a format suitable for your analysis

First set the working directory to the desktop, where the data is.
Load relevant libraries.
And read them. Use ymd from lubridate to modify the format of the date.
```{r}
setwd("~/Desktop")
library(dplyr)
library(lubridate)
act_file <- read.csv("activity.csv")
act_file$date <- ymd(act_file$date)
act_file_clean <- act_file[complete.cases(act_file),]
```

###Question 2. What is mean total number of steps taken per day?
histogram of the total number of steps taken each day
```{r}
sub1 <- group_by(act_file_clean[,1:2], date) %>% 
  summarize(totalsteps = sum(steps))
png("plot1.png", width = 480, height = 480)
with(sub1, barplot(totalsteps))
dev.off()

with(sub1, barplot(totalsteps))
```

Calculate and report the mean and median total number of steps taken per day
```{r}
sub2<- group_by(act_file_clean[,1:2], date) %>% 
  summarize(mean = mean(steps), median = median(steps))
sub2
```

###Question 3. What is the average daily activity pattern?
Time series plot of 5-min interval and avg number of steps by days
```{r}
png("plot2.png", width = 480, height = 480)
with(sub2, plot(date,mean,type="l"))
dev.off()
with(sub2, plot(date,mean,type="l"))
```
The interval having the max steps
```{r}
sub3 <- group_by(act_file_clean, interval) %>% 
  summarize(steps=max(steps)) 
maxinterval <- sub3 %>% 
  filter(steps==max(sub3$steps)) %>% 
  select(interval)
maxinterval
```

###Question 4. Imputing NAs
```{r}
nafile <- act_file[!complete.cases(act_file),]
nafile$steps <- 0
act_imputing <- rbind(nafile, act_file_clean)
```
Histogram with imputed figures.
```{r}
sub7 <- group_by(act_imputing[,1:2], date) %>% 
  summarize(totalsteps = sum(steps))
png("plot3.png", width = 480, height = 480)
with(sub7, barplot(totalsteps))
dev.off()

with(sub7, barplot(totalsteps))
```
mean and median with the imputed
```{r}
sub7_2<- group_by(act_imputing[,1:2], date) %>% 
  summarize(mean = mean(steps), median = median(steps))

sub7_2
```
 ###Question 5. panel plot
```{r}
library(lattice)
act_file_clean$wday <- wday(act_file_clean$date)
act_file_clean<-act_file_clean %>% 
  mutate(wd = wday>=6)
sub8 <- act_file_clean[,c(1,3,5)]
wkdy <- sub8 %>% 
  filter(wd=="FALSE") %>%
  select(steps, interval) %>% 
  group_by(interval) %>% 
  summarize(steps=mean(steps)) %>%
  mutate(id = "weekday")
wked <- sub8 %>% 
  filter(wd=="TRUE") %>%
  select(steps, interval) %>% 
  group_by(interval) %>% 
  summarize(steps=mean(steps)) %>% 
  mutate(id = "weekend")
sub8 <- rbind(wkdy, wked)
png("plot4.png", width = 480, height = 480)
xyplot(steps ~ interval | id, data = sub8, layout = c(1, 2), type= "l")
dev.off()

xyplot(steps ~ interval | id, data = sub8, layout = c(1, 2), type= "l")
```







